---
title: "QAQC Workshop"
author: "Alexandra Lalor"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages
library(tidyverse)
library(writexl)
library(dplyr)
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
getwd()

# Load in data. Change file path based on user name!
path_data <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/data/"
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/output/"
```

### Load Data

```{r}
# Load in data
Fuels1000_all <- read.csv(paste0(path_data, "GRCA_Surface Fuels - 1000Hr_XPT.csv"))
```

```{r}
# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA

# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day")) %>% 
  # Separate MacroPlotName to get Monitoring type
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited), Year == 2024)

# Filter for only headers
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited), Year == 2024)
```

### Create blank data frame

```{r}
# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "",
                           "MacroPlot.Name" = "",
                           "Date" = "",
                           "Error" = "No Error", 
                           "Comment" = "")
```

# FUNCTIONS

```{r}
# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  # Identify errors 
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%   
    # Data is equal exactly to a value
    filter((values_check) %>% replace_na(TRUE)) %>%   
    select("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment")
  
  # If no errors, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}
```

# PROTOCOL - SURFACE FUELS

## 1000HR CWD

This code conducts quality control checks on surface fuels data within the surface fuels coarse woody debris (1000-hr) data set.

It checks for: header info (number of transects, length of transects), decay class values, fuel constants, duplicate values, transect numbers, reasonable slope numbers, reasonable diameter values.

### Fuel 1000 Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header NumTran"
query_message <- "Number of Transects" 
values_data <- data$NumTran
values_valid <- 4
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that North Rim TranLen = 50 (or very close). North Rim plots include PIAB, PIEN, and PIPN.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MonitoringType %in% c("PIAB", "PIEN", "PIPN"))
query <- "Fuel 1000 Header NRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 50  
values_check <-!values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_N <- qaqc(data, query, query_message, values_check)
```

-   Check that S Rim TranLen = 100 (or very close). South Rim plots include PIED and PIPO.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MonitoringType %in% c("PIED", "PIPO"))
query <- "Fuel 1000 Header SRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 100 
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_S <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline}

-   Check that DecayCl = 3 or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 DecayClass"
query_message <- "Decay Class"
values_data <- data$DecayCl
values_valid <- c(3,4)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_DecayCl <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that CWDFuConSt = Monitoring Type

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 FuelConst"
query_message <- "Fuel Constant"
values_data <- data$CWDFuConSt
values_valid <- data$MonitoringType
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- (!values_data %in% values_valid) %>% replace_na(TRUE)

# Identify errors
errors_Fuels1000_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_Rosner.Rmd

#### Diameter

# Filter all data to remove NA values for outlier test
data_temp <- Fuels1000_all %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(Dia))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$Dia)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Dia <- min(outliers) - 1

# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Dia"
query_message <- "Diameter"
values_data <- data$Dia
values_valid <- seq(3, max_Dia, by = 0.5)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_Dia <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Errors

```{r}
# Save to master error list
errors_Fuels1000 <- rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen_N, errors_Fuels1000_TranLen_S, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)

# Table of results for quick check
kable(errors_Fuels1000, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors_Fuels1000, paste0(path_errors, "errors_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_Fuels1000.xlsx"))
```
