---
title: "QAQC Workshop - Fuels Fine"
author: "FFI + R Working Group"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages

# tidyverse and dplyr have lots of useful functions for data cleaning
library(tidyverse)
library(dplyr)
# writexl is used to create excel files from within R
library(writexl)
# knitr is used to create output files from R Markdown
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory
getwd()

# Load in data. Change file path based on user name!
path_data <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/data/"
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/output/"
```

### Load Data

```{r}
# Load in data
FuelsFine_all <- read.csv(paste0(path_data, "GRCA_Surface Fuels - Fine_XPT.csv"))
```

```{r}
# Clean up data frame
FuelsFine_all <- FuelsFine_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day")) %>% 
  # Separate MacroPlotName to get Monitoring type
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Ensure blanks in Visited column are NA
FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA

# Filter for only data
FuelsFine_data <- FuelsFine_all %>%
  filter(is.na(Visited), Year == 2024)

# Filter for only headers
FuelsFine_header <- FuelsFine_all %>%
  filter(!is.na(Visited), Year == 2024)
```

### Create blank data frame

```{r}
# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "",
                           "MacroPlot.Name" = "",
                           "Date" = "",
                           "Error" = "No Error", 
                           "Comment" = "")
```

### Create Functions

```{r}
# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  # Identify errors 
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%
    filter(!(values_check) %>% replace_na(TRUE)) %>%   
    select("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment")
  
  # If no errors, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}
```

## FINE FUELS

This code conducts quality control checks on surface fuels data within the surface fuels fine (FWD) data set.

It checks for: header info (number of transects, length of transects for One/Ten/Hundred Hour), fuel constants, transect numbers, reasonable azimuth numbers, reasonable slope numbers, reasonable One/Ten/Hundred Hour count values.

### Fuel Fine Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that OneHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header OneHrTranLen" 
query_message <- "One Hour Transect Length"  
values_data <- data$OneHrTranLen
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_OneHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that TenHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header TenHrTranLen" 
query_message <- "Ten Hour Transect Length"  
values_data <- data$TenHrTranLen 
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_TenHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that HunHrTranLen = 12

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header HunHrTranLen" 
query_message <- "Hundrend Hour Transect Length"  
values_data <- data$HunHrTranLen 
values_valid <- 12
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_HunHrTranLen <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine FuelConst

[Problem:]{.underline} Incorrect fine woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that FWDFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$FWDFuConSt 
values_valid <- data$MonitoringType  
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Hits

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 1 sample point per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 4

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp <- FuelsFine_data %>%
  filter(!is.na(Transect)) %>% 
  group_by(MacroPlot.Name) %>% 
  summarize("sum_hits" = n())

# Set parameters 
data <- merge(FuelsFine_data, data_temp, all.x = T)
query <- "Fuel Fine Hits" 
query_message <- "Number of Hits" 
values_data <- data$sum_hits
values_valid <- 4
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Values

[Problem:]{.underline}  Values for transect, azimuth, slope, and counts for One, Ten, Hundred hour fuels are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Azimuth values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Azimuth"
query_message <- "Azimuth"
values_data <- data$Azimuth
values_valid <- seq(0, 359, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Azimuth <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that One Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(OneHr))
test <- rosnerTest(data_temp$OneHr)
outliers <- test$all.stats$Value[test$all.stats$Outlier]
max_One <- min(outliers) - 1

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values OneHr" 
query_message <- "One Hour Count Outlier" 
values_data <- data$OneHr
values_valid <- seq(0, max_One, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_OneHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Ten Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(TenHr))
test <- rosnerTest(data_temp$TenHr)
outliers <- test$all.stats$Value[test$all.stats$Outlier]
max_Ten <- min(outliers) - 1

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values TenHr" 
query_message <- "Ten Hour Count Outlier" 
values_data <- data$TenHr
values_valid <- seq(0, max_Ten, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TenHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Hundred Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(HunHr))
test <- rosnerTest(data_temp$HunHr)
outliers <- test$all.stats$Value[test$all.stats$Outlier]
max_Hun <- min(outliers) - 1

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values HunHr" 
query_message <- "Hundred Hour Count Outlier" 
values_data <- data$HunHr
values_valid <- seq(0, max_Hun, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_HunHr <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Errors

```{r}
# Save to master error list 
errors_FuelsFine <- unique(rbind(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen, errors_FuelsFine_Hits, errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr))
```

```{r}
remove(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen, errors_FuelsFine_Hits, errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr)
```

```{r}
# Table of results for quick check
kable(errors_FuelsFine, "pipe")
```

```{r}
# Save as CSV or XLSX 
path_errors
write_xlsx(errors_FuelsFine, paste0(path_errors, "errors_FuelsFine.xlsx"))
```
