---
title: "QAQC Workshop - Fuels Duff Litt"
author: "FFI + R Working Group"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages

# tidyverse and dplyr have lots of useful functions for data cleaning
library(tidyverse)
library(dplyr)
# writexl is used to create excel files from within R
library(writexl)
# knitr is used to create output files from R Markdown
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory
getwd()

# Load in data. Change file path based on user name!
path_data <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/data/"
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/Workshop1_QAQC/output/"
```

### Load Data

```{r}
# Load in data
FuelsDuffLitt_all <- read.csv(paste0(path_data, "GRCA_Surface Fuels - Duff_Litter_XPT.csv"))
```

```{r}
# Clean up data frame
FuelsDuffLitt_all <- FuelsDuffLitt_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day")) %>% 
  # Separate MacroPlotName to get Monitoring type
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Ensure blanks in Visited column are NA
FuelsDuffLitt_all$Visited[FuelsDuffLitt_all$Visited==""] <- NA

# Filter for only data
FuelsDuffLitt_data <- FuelsDuffLitt_all %>%
  filter(is.na(Visited), Year == 2024)

# Filter for only headers
FuelsDuffLitt_header <- FuelsDuffLitt_all %>%
  filter(!is.na(Visited), Year == 2024)
```

### Create blank data frame

```{r}
# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "",
                           "MacroPlot.Name" = "",
                           "Date" = "",
                           "Error" = "No Error", 
                           "Comment" = "")
```

### Create Functions

```{r}
# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  # Identify errors 
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%
    filter(!(values_check) %>% replace_na(TRUE)) %>%   
    select("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment")
  
  # If no errors, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}
```

## DUFF LITTER

This code conducts quality control checks on surface fuels data within the surface fuels duff litter data set.

It checks for: number of transects, fuel constants, number of points sampled, reasonable litter values, reasonable duff values.

### Fuel DL Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsDuffLitt_header
query <- "Fuel DL Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_NumTran <- qaqc(data, query, query_message, values_check)
```

### Fuel DL FuelConst

[Problem:]{.underline} Incorrect duff & litter fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that DLFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$DLFuConSt 
values_valid <- data$MonitoringType  
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Hits

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 10 sample points per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 40

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp <- FuelsDuffLitt_data %>%
  filter(!is.na(SampLoc)) %>% 
  group_by(MacroPlot.Name) %>% 
  summarize("sum_hits" = n())

# Set parameters 
data <- merge(FuelsDuffLitt_data, data_temp, all.x = T)
query <- "Fuel DL Hits" 
query_message <- "Number of Hits" 
values_data <- data$sum_hits
values_valid <- 40
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Values

[Problem:]{.underline}  Values for litter and duff depth are missing or unreasonable.

[Procedure:]{.underline}

-   Sort by LittDep. Check that litter depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(LittDep))
data_rosner <- rosnerTest(data_temp$LittDep)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Litt <- min(outliers) - 0.1

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values LittDep" 
query_message <- "Litter Depth Outlier" 
values_data <- data$LittDep
values_valid <- max_Litt
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_LittDep <- qaqc(data, query, query_message, values_check)
```

-   Sort by DuffDep. Check that duff depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(DuffDep))
data_rosner <- rosnerTest(data_temp$DuffDep)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Duff <- min(outliers) - 1

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values DuffDep" 
query_message <- "Duff Depth Outlier" 
values_data <- data$DuffDep
values_valid <- max_Duff
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_DuffDep <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Errors

```{r}
# Save to master error list 
errors_FuelsDuffLitt <- unique(rbind(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep))
```

```{r}
remove(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep)
```

```{r}
# Table of results for quick check
kable(errors_FuelsDuffLitt, "pipe")
```

```{r}
# Save as XLSX
path_errors
write_xlsx(errors_FuelsDuffLitt, paste0(path_errors, "errors_FuelsDuffLitt.xlsx"))
```
